---
output:
  pdf_document: default
  html_document: default
---

```{r, include=FALSE}
library(tidyverse)
library(Rmisc)
library(ggpubr)
require(dplyr)
library(forcats)
library(RColorBrewer)

library(readr)
library(magrittr)
library(gridExtra)
library(vtable)
library(kableExtra)
source("~/scripts/exportPlot.R")

```

```{r data, include=FALSE}
##read in data

setwd("~/data/gvaginalis_expEvo/phenotype-plots/")

##Aerobic growth curves
gcA <- read_csv("2022.05.09_GV14018_GCs_biofilm_ancestor_all P5_all P10_JM.csv")

gv2 <- readRDS("~/data/gvaginalis_expEvo/2022.06.16_gv_growthcurves/GV2-aerobic-data")
gv2.anc <- filter(gv2, Sample == "GV2 ancestor ")

gv.14 <- readRDS("~/data/gvaginalis_expEvo/2022.06.16_gv_growthcurves/GV14018-aerobic-data")
gv.14.anc <- filter(gv.14, Sample == "GV14018 ancestor ")
gv.anc <- rbind(gv2.anc, gv.14.anc)
gv.anc <- na.omit(gv.anc)

gv.both <- rbind(gv2, gv.14)
gv.both <- na.omit(gv.both)


##Anaerobic growth curves
gcAn <- read_csv("2022-11-11-GV2-14018-anaerobicGrowthCurves.csv", col_names = TRUE)

##biofilm and CV quantification
biofilmQuant<- read.csv("2022.12.02_area scanning_GV2 and 14018.csv", header=T) %>%
  mutate(Strain = as.factor(Strain))

planktonicQuant <- read.csv("2023.02.15_GV planktonic evolved biofilms_no shaking_all strains.csv", header=T)

##AutoAggregation
sediment_all <- read.csv("2024.03.26_GV sedimentation_all strains_manuscript edits.csv", header=T)

```

```{r colors, include =FALSE}
#"#FFF5F0" "#FEE0D2" "#FCBBA1" "#FC9272" "#FB6A4A" "#EF3B2C" "#CB181D" "#A50F15" "#67000D"

GV2_colors <- c( "#0A529C", "#9ECAE1")
GV14018_colors <- c("#A50F15", "#F4BEBA")
Anc_colors <- c("#A50F15", "#0A529C" )

boxplotLine <- "grey60"
```





# OD & CV Biofilm quantification

### Boxplots by strain Evolved vs. Ancestor

```{r, echo = FALSE}

GV2 <- filter(biofilmQuant, Strain == "GV2")
GV14018 <- filter(biofilmQuant, Strain == "14018")


#OD_biofilm
GV2_biofilm <- ggplot(data=GV2, aes(x=Type, y=OD.biofilm, fill=Type)) + 
  geom_boxplot() + 
  xlab("GV strain") + 
  ylab("OD(600) biofilm") + 
  ggtitle("") + 
  coord_cartesian(ylim=c(0.05,0.3)) + 
  theme_minimal() + 
  scale_fill_manual(values = GV2_colors) +
  theme(plot.title = element_text(hjust = 0.5), text=element_text(size= 12), legend.title = element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  stat_compare_means(method = "t.test", label = "p.signif", ref.group = "Ancestor")

GV2_biofilm

#ExportPlot(GV2_biofilm, "phenotpyePlots_04.12.24/boxplots/GV2_biofilm", height= 5, width= 7)

GV14018_biofilm <- ggplot(data=GV14018, aes(x=Type, y=OD.biofilm, fill=Type)) + 
  geom_boxplot() + 
  xlab("GV strain") + 
  ylab("OD(600) biofilm") + 
  ggtitle("") + 
  coord_cartesian(ylim=c(0.05,0.3)) + 
  theme_minimal() + 
  scale_fill_manual(values= GV14018_colors) +
  theme(plot.title = element_text(hjust = 0.5), text=element_text(size= 12), legend.title = element_blank(), axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) + 
  stat_compare_means(method = "t.test",  label = "p.signif", ref.group = "Ancestor")

GV14018_biofilm

#ExportPlot(GV14018_biofilm, "phenotpyePlots_04.12.24/boxplots/GV14018_biofilm", height=5, width =7)


#OD_CV
GV2_CV <-  ggplot(GV2, aes(x=Type, y=CV, fill=Type)) + 
  geom_boxplot() + 
  xlab("GV strain") + 
  ylab("OD(570) crystal violet") + 
  ggtitle("") + 
  coord_cartesian(ylim=c(0.05,3.2)) + 
  theme_minimal() + 
  scale_fill_manual(values= GV2_colors) +
  theme(plot.title = element_text(hjust = 0.5), text=element_text(size= 12), axis.title.x=element_blank(), axis.text.x=element_blank(),axis.ticks.x=element_blank(), legend.position = "none") + 
  stat_compare_means(method = "t.test", label = "p.signif", ref.group = "Ancestor")


GV2_CV

#ExportPlot(GV2_CV, "phenotpyePlots_04.12.24/boxplots/GV2_CV", height = 5, width = 7)

GV14018_CV <- ggplot(data=GV14018, aes(x=Type, y=CV, fill=Type)) + 
  geom_boxplot() + 
  xlab("GV strain") + 
  ylab("OD(570) crystal violet") + 
  ggtitle("") + 
  coord_cartesian(ylim=c(0.05,3.2)) + 
  theme_minimal() + 
  scale_fill_manual(values= GV14018_colors) +
  theme(plot.title = element_text(hjust = 0.5), text=element_text(size= 12), axis.title.x=element_blank(), axis.text.x=element_blank(),axis.ticks.x=element_blank(), legend.position = "none") + 
  stat_compare_means(method = "t.test", label = "p.signif", ref.group = "Ancestor")

GV14018_CV

#ExportPlot(GV14018_CV, "phenotpyePlots_04.12.24/boxplots/GV14018_CV", height = 5, width = 7)


GV2Arranged <- ggarrange(GV2_biofilm, GV2_CV, nrow = 2, ncol =1, common.legend = TRUE)
GV14018Arranged <- ggarrange(GV14018_biofilm, GV14018_CV, nrow = 2, ncol =1, common.legend = TRUE)

ExportPlot(GV2Arranged, "phenotpyePlots_04.12.24/boxplots/GV2_CV_BF", height = 6, width = 7)

ExportPlot(GV14018Arranged, "phenotpyePlots_04.12.24/boxplots/GV14018_CV_BF", height = 6, width = 7)

```

### OD & CV of 14018 Ancestor vs. GV2 Ancestor

```{r, echo = FALSE}
Ancestors <- filter(biofilmQuant, Type == "Ancestor")

Ancestors_biofilm <- ggplot(data=Ancestors, aes(x=Strain, y=OD.biofilm, fill=Strain)) + 
  geom_boxplot() + 
  xlab("GV strain") +
  ylab("OD(600) biofilm") + 
  ggtitle("") + 
  scale_fill_manual(values= Anc_colors) + 
  coord_cartesian(ylim=c(0.05,0.3)) + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(text=element_text(size= 12)) + 
  theme(axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(), legend.title = element_blank()) + 
   stat_compare_means(method = "t.test", label = "p.signif", ref.group = "14018")

Ancestors_biofilm


#ExportPlot(Ancestors_biofilm, "phenotpyePlots_04.12.24/boxplots/Ancestors_biofilm", height=5, width=7)


Ancestors_CV <- ggplot(data=Ancestors, aes(x=as.factor(Strain), y=CV, fill=Strain)) + 
  geom_boxplot() + 
  xlab("GV strain") + 
  ylab("OD(570) crystal violet") + 
  ggtitle("") + 
  scale_fill_manual(values= Anc_colors) + 
  coord_cartesian(ylim=c(0.05,3)) + 
  theme_minimal() + 
  theme(text=element_text(size= 12)) + 
  theme(axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(), legend.title = element_blank()) + 
  stat_compare_means(method = "t.test", label = "p.signif", ref.group = "14018")

Ancestors_CV

AncestorsArranged <- ggarrange(Ancestors_biofilm, Ancestors_CV, ncol = 2, nrow = 1, common.legend = TRUE)

ExportPlot(AncestorsArranged, "phenotpyePlots_04.12.24/boxplots/Ancestors_Arranged", height=5, width = 8)
```
#stats to look at biofilm density between ancestor and evolved (mann U test) 
```{r}
#GV2 OD biofilm
GV2_biofilm_stats <- wilcox.test(OD.biofilm ~ Type, data = GV2,
                   exact = FALSE, alternative = "less")
#GV2 CV
GV2_CV_stats <- wilcox.test(CV ~ Type, data = GV2,
                   exact = FALSE, alternative = "less")

#GV14018 OD biofilm
GV14018_biofilm_stats <- wilcox.test(OD.biofilm ~ Type, data = GV14018,
                   exact = FALSE, alternative = "less")

#GV14018 CV
GV14018_CV_stats <- wilcox.test(CV ~ Type, data = GV14018,
                   exact = FALSE, alternative = "less")
```


```{r}
#GV2 vs. GV14018
Ancestor_biofilm_stats <- wilcox.test(OD.biofilm ~ Sample, data=Ancestors)
Ancestor_CV_stats <- wilcox.test(CV ~ Sample, data=Ancestors_CV)

print(GV2_biofilm_stats)
print(GV2_CV_stats)

print(GV14018_biofilm_stats)
print(GV14018_CV_stats)
```


# Planktonic evolved biofilm quantification - shaking
```{r}
planktonicQuant <- planktonicQuant %>%
  mutate(Type = case_when(grepl("ancestor", Sample) ~ "Ancestor", TRUE ~ "Evolved"))

GV2_plank <- filter(planktonicQuant, Strain == "GV2")
GV14018_plank <- filter(planktonicQuant, Strain == "GV14018")
```

#GV2 graphs
```{r}
GV2_biofilm_plank <- ggplot(data=GV2_plank, aes(x=as.factor(Type), y=OD_biofilm, fill= Type)) + 
  geom_boxplot() + 
  xlab("GV strain") + 
  ylab("OD(600) biofilm") + 
  ggtitle("") + 
  coord_cartesian(ylim=c(0.05,0.3)) + 
  theme_minimal() + theme(text=element_text(size= 12)) + theme(axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank()) + 
  scale_fill_manual(values = GV2_colors, name = "") +
  stat_compare_means(method = "t.test", label = "p.signif", ref.group = "Ancestor")


print(GV2_biofilm_plank)

GV2_CV_plank <- ggplot(data=GV2_plank, aes(x=as.factor(Type), y=OD_CV, fill=Type)) + 
  geom_boxplot() + 
  xlab("GV strain") + 
  ylab("OD(570) crystal violet") + 
  ggtitle("") + 
  coord_cartesian(ylim=c(0.05,3.2)) + 
  theme_minimal() + 
  theme(text=element_text(size= 12)) + 
  theme(axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank()) + 
  scale_fill_manual(values = GV2_colors, name = "") +
  stat_compare_means(method = "t.test", label = "p.signif", ref.group = "Ancestor")

print(GV2_CV_plank)

plank_arrange <- ggarrange(GV2_biofilm_plank, GV2_CV_plank, ncol = 1, nrow = 2, common.legend = TRUE)
ExportPlot(plank_arrange, "phenotpyePlots_04.12.24/GV2-plank-biofilms", height = 6, width = 6)
```

#14018 graphs
```{r}
GV14018_biofilm_plank <- ggplot(data=GV14018_plank, aes(x=as.factor(Type), y=OD_biofilm, fill = Type)) + 
  geom_boxplot() + 
  xlab("GV strain") +
  ylab("") + 
  ggtitle("") + 
  coord_cartesian(ylim=c(0.05,0.3)) + 
  theme_minimal() + 
  theme(text=element_text(size= 12)) + 
  theme(axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank()) + 
  scale_fill_manual(name= "", values = GV14018_colors) +
  stat_compare_means(method = "t.test", label = "p.signif", ref.group = "Ancestor")


print(GV14018_biofilm_plank)

GV14018_CV_plank <- ggplot(data=GV14018_plank, aes(x=as.factor(Type), y=OD_CV, fill=Type)) + 
  geom_boxplot() + 
  xlab("GV strain") + 
  ylab("OD(570) crystal violet") + 
  ggtitle("") + 
  coord_cartesian(ylim=c(0.05, 3.2)) +
  theme_minimal() + 
  theme(text=element_text(size= 12)) + 
  theme(axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank()) + 
  scale_fill_manual(name= "", values = GV14018_colors) +
  stat_compare_means(method = "t.test", label = "p.signif", ref.group = "Ancestor")


print(GV14018_CV_plank)

GV14018_plankArrange <- ggarrange(GV14018_biofilm_plank, GV14018_CV_plank, ncol = 1, nrow = 2, common.legend = TRUE)
ExportPlot(GV14018_plankArrange, "phenotpyePlots_04.12.24/GV14018-plank-biofilms", height = 6, width = 6)

```


```{r}
#ExportPlot(GV2_biofilm_plank, "phenotpyePlots_04.12.24/boxplots/GV2_biofilm_plank", height = 5, width = 7)
#ExportPlot(GV2_CV_plank, "phenotpyePlots_04.12.24/boxplots/GV2_CV_plank", height = 5, width = 7)
#ExportPlot(GV14018_biofilm_plank, "phenotpyePlots_04.12.24/boxplots/GV14018_biofilm_plank", height = 5, width = 7)
#ExportPlot(GV14018_CV_plank, "phenotpyePlots_04.12.24/boxplots/GV14018_CV_plank", height = 5, width = 7)
```
### 

# Growth curves

## Aerobic GV14018 Ancestor vs. GV2 Ancestor

```{r, echo = FALSE}
gv.anc$Value <- as.numeric(gv.anc$Value)
gvs <- summarySE(gv.anc, measurevar="Value", groupvars=c("Sample","Time"))


plot <- ggplot(gvs, aes(x=Time, y=Value, colour=Sample))  + 
  geom_line(aes(group=Sample)) + 
  theme_classic() + 
  geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.001) +   scale_x_discrete(breaks=c("00:00:00","24:00:00","48:00:00"), labels=c("0", "24", "48")) + 
  ylim(0.1, 1) + ylab("OD600") + 
  scale_color_manual(name= "Population",labels= c("GV14018 ancestor", "GV2 ancestor"), values = Anc_colors) +
  theme(legend.title = element_blank()) + 
  ggtitle("Aerobic")

plot

ExportPlot(plot, "phenotpyePlots_04.12.24/Aerobic_GCs/Ancestor_gc", height = 5, width = 7)

```

### Aerobic Ancestor vs. Evolved

```{r, echo = FALSE}
##making sure the class of each column is correct
gv.both <- gv.both %>% 
  mutate(Value = as.numeric(Value), Sample = as.character(Sample), 
         Time = as.factor(Time), 
         Replicate = as.factor(Replicate))

##apply function
tgc <- summarySE(gv.both, measurevar="Value", groupvars=c("Sample","Time")) %>%
  mutate(Type = case_when(grepl("ancestor", Sample) ~ "Ancestor", TRUE ~"Evolved"))
  

##filter so each Experiment is separated out to plot
cfs1 <- filter(tgc, Sample %in% c("GV14018 ancestor ", "2021-10-01-GV14018-CFS-1_P10 ", "2021-10-01-GV14018-CFS-2_P10 ", "2021-10-01-GV14018-CFS-3_P10 "))

nil1  <- filter(tgc, Sample %in% c("GV14018 ancestor ", "2021-10-01-GV14018-Nil-1_P10 ", "2021-10-01-GV14018-Nil-2_P10 ","2021-10-01-GV14018-Nil-3_P10 "))

##GV14018 Nil population
nil1_plot <- ggplot(nil1, aes(x=Time, y=Value, colour=Type))  + 
  geom_line(aes(group=Sample)) + 
  theme_classic() + 
  scale_color_manual(name= "Population", values= GV14018_colors) +
  geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.001) + 
  theme(legend.title = element_blank()) +
  ylim(0.1, 1.2) +
  ylab("OD600")  + 
  scale_x_discrete(breaks=c("00:00:00","24:00:00","48:00:00"), labels=c("0", "24", "48"))
nil1_plot

ExportPlot(nil1_plot, "phenotpyePlots_04.12.24/Aerobic_GCs/GV14018_nil_GC", height = 5, width = 7)

##GV14018 CFS population
cfs1_plot <- ggplot(cfs1, aes(x=Time, y=Value, colour=Type))  + 
  geom_line(aes(group=Sample)) + 
  theme_classic() + 
  scale_color_manual(name= "Population", values= GV14018_colors) +
  geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.001) + 
  ylim(0.1, 1.2) +
  theme(legend.title = element_blank()) +
  ylab("OD600")  + 
  scale_x_discrete(breaks=c("00:00:00","24:00:00","48:00:00"), labels=c("0", "24", "48"))
cfs1_plot
ExportPlot(cfs1_plot, "phenotpyePlots_04.12.24/Aerobic_GCs/GV14018_cfs_GC", height = 5, width = 7)


```

# Aerobic growth curves : GV2

```{r, echo = FALSE}
tgc2 <- summarySE(gv.both, measurevar= "Value", groupvars=c("Sample", "Time"))  %>%
  mutate(Type = case_when(grepl("ancestor", Sample) ~ "Ancestor", TRUE ~"Evolved"))

##filter so each Experiment is separated out to plot
tgc_jun <- filter(tgc2,   Sample %in% c("GV2 ancestor ","Experiment 1_Population 1 ", "Experiment 1_Population 2 ", "Experiment 1_Population 3 "))
tgc_aug <- filter(tgc2,  Sample %in% c("GV2 ancestor ", "Experiment 2_Population 1 ", "Experiment 2_Population 2 ", "Experiment 2_Population 3 "))
tgc_nil <- filter(tgc2,   Sample %in% 
                    c("GV2 ancestor ", "Experiment 3_Population 1_end ", "Experiment 3_Population 2_end ", "Experiment 3_Population 3_end "))


##make plots
june_plot <- ggplot(tgc_jun, aes(x=Time, y=Value, colour=Type))  + 
  geom_line(aes(group=Sample)) + 
  theme_classic() + 
  geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.001) + 
  theme(legend.title = element_blank()) +
  ylim(0.1, 1) + 
  ylab("OD600") + 
  scale_color_manual(name= "Population", values= GV2_colors) +
  scale_x_discrete(breaks=c("00:00:00","24:00:00","48:00:00"), labels=c("0", "24", "48"))

aug_plot <- ggplot(tgc_aug, aes(x=Time, y=Value, colour=Type))  + 
  geom_line(aes(group=Sample)) + 
  theme_classic() + 
  geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.001) + 
  ylim(0.1, 1 ) + 
  ylab("OD600") +
  theme(legend.title = element_blank()) +
  scale_color_manual(name= "Population", values= GV2_colors) + scale_x_discrete(breaks=c("00:00:00","24:00:00","48:00:00"), labels=c("0", "24", "48"))

nil_plot <- ggplot(tgc_nil, aes(x=Time, y=Value, colour=Type))  + 
  geom_line(aes(group=Sample)) + 
  theme_classic() + 
  scale_color_manual(name= "Population", values= GV2_colors) + 
  geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.001) + ylim(0.1, 1) + ylab("OD600")  + 
  theme(legend.title = element_blank()) +
  scale_x_discrete(breaks=c("00:00:00","24:00:00","48:00:00"), labels=c("0", "24", "48"))

nil_plot

##view plots
ExportPlot(june_plot, "phenotpyePlots_04.12.24/Aerobic_GCs/GV2_June_CFS", height = 5, width = 7)
ExportPlot(aug_plot, "phenotpyePlots_04.12.24/Aerobic_GCs/GV2_Aug_CFS", height = 5, width = 7)
ExportPlot(nil_plot, "phenotpyePlots_04.12.24/Aerobic_GCs/GV2_Nil", height = 5, width = 7)
```

# GV14018 Aerobic

```{r, echo=FALSE}

##filter so each Experiment is separated out to plot
cfs1 <- filter(tgc2, Sample %in% c("GV14018 ancestor ", "2021-10-01-GV14018-CFS-1_P10 ", "2021-10-01-GV14018-CFS-2_P10 ", "2021-10-01-GV14018-CFS-3_P10 "))

nil1 <- filter(tgc2, Sample %in% c("GV14018 ancestor ", "2021-10-01-GV14018-Nil-1_P10 ", "2021-10-01-GV14018-Nil-2_P10 ","2021-10-01-GV14018-Nil-3_P10 "))

##make plots
cfsPlot <- ggplot(cfs1, aes(x=Time, y=Value, colour=Type))  + geom_line(aes(group=Sample)) + theme_classic() + 
  scale_color_manual(name= "Population", values= GV14018_colors) +
  geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.001) + 
  ylim(0.1, 1) + 
  ylab("OD600")  + 
  theme(legend.text = element_blank()) +
  scale_x_discrete(breaks=c("00:00:00","24:00:00","48:00:00"), labels=c("0", "24", "48"))

nilPlot <- ggplot(nil1, aes(x=Time, y=Value, colour=Type))  + geom_line(aes(group=Sample)) + theme_classic() + 
  scale_color_manual(name= "Population", values= GV14018_colors) +
  geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.001) + ylim(0.1, 1) + ylab("OD600")  + 
  theme(legend.text = element_blank()) +
  scale_x_discrete(breaks=c("00:00:00","24:00:00","48:00:00"), labels=c("0", "24", "48"))

##view plots
ExportPlot(nilPlot, "phenotpyePlots_04.12.24/Aerobic_GCs/GV14018_nil", height= 5, width=7)
ExportPlot(cfsPlot, "phenotpyePlots_04.12.24/Aerobic_GCs/GV14018_cfs", height = 5, width = 7)

```

# Anaerobic growth curves

```{r, echo = FALSE}
##Anaerobic growth curves
gc <- read_csv("2023.01.10_14018 GCs_anaerobic_ancestor redone.csv", col_names = TRUE)
```

```{r, echo= FALSE}
gc <- na.omit(gc)
##format so Time is a column
rownames(gc) <- gc$Time
gc <- t(gc)
gc <- gc[-1,]
gc <- as.data.frame(gc)
gc <- tibble::rownames_to_column(gc, "Sample")
gc <- gc %>% separate(Sample, sep= "[#]", into= c("Sample", "Replicate"))

##Remove extra characters from the Replicate column
gc$Replicate <- gsub( ")", "", gc$Replicate)
gc$Sample <- sub("\\_$", "", gc$Sample)
name1 <- gc[,-c(1,2)]
list1 <- paste0(colnames(name1))
list1 <- c(list1)

##gather so that each timepoint has a separate Time value (makes it easier to plot)
lgc_onlynamed <- gather(gc, key= 'Time', value= 'Value',  colnames(name1))
lgc_onlynamed$Time <- as.factor(lgc_onlynamed$Time)
data <- lgc_onlynamed

##format data so all the columns are the correct class
data$Sample <- as.character(data$Sample)
data$Time <- as.factor(data$Time)
data$Replicate <- as.factor(data$Replicate)
data$Value <- as.numeric(data$Value)
##apply function

tgc3  <- summarySE(data, measurevar= "Value", groupvars=c("Sample", "Time")) %>%
  mutate(Type = case_when(grepl("ancestor", Sample) ~ "Ancestor", TRUE ~ "Evolved"))


tgc_ordered <- tgc3 %>% 
  mutate(Time=factor(Time)) %>% 
  mutate(Time=fct_relevel(Time,c("0:00:00", "0:30:00", "1:00:00", "1:30:00", "2:00:00", "2:30:00", "3:00:00", "3:30:00", "4:00:00", "4:30:00", "5:00:00", "5:30:00", "6:00:00", "6:30:00", "7:00:00", "7:30:00", "8:00:00", "8:30:00", "9:00:00", "9:30:00"))) %>% 
  arrange(Time)
```

# GV14018 anaerobic

```{r, echo = FALSE}
tgc_oct_cfs <- filter(tgc_ordered,   Sample %in% c("GV14018_ancestor_2" , "14018_Oct_CFS_1_P10", "14018_Oct_CFS_2_P10" , "14018_Oct_CFS_3_P10" ))
tgc_oct_nil <- filter(tgc_ordered,   Sample %in% c("GV14018_ancestor_2" , "14018_Oct_Nil_1_P10", "14018_Oct_Nil_2_P10" , "14018_Oct_Nil_3_P10" ))

anc_comparison <- filter(tgc_ordered,   Sample %in% c("GV2_ancestor" , "GV14018_ancestor_2")) 


##make plots
GV14018_cfs <- ggplot(tgc_oct_cfs, aes(x=Time, y=Value, color=Type))  + 
  geom_line(aes(group=Sample)) + 
  theme_classic() + 
  scale_color_manual(values= GV14018_colors) + 
  geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.001) + 
  ylim(0.1, 1.2) + 
  ylab("OD600")  + 
  scale_x_discrete(breaks=c("00:00:00","24:00:00","48:00:00"), labels=c("0", "24", "48")) +
  theme(legend.title= element_blank()) + 
  ggtitle("14018 anaerobic curves: October_CFS vs. Ancestor")

print(GV14018_cfs)

ExportPlot(GV14018_cfs, "phenotpyePlots_04.12.24/AnAerobic_GCs/GV14018_cfs", height = 5, width = 7)


GV14018_nil <- ggplot(tgc_oct_nil, aes(x=Time, y=Value, colour=Type))  + 
  geom_line(aes(group=Sample)) + 
  theme_classic() + 
  scale_color_manual(name= "Population", values= GV14018_colors) + 
  geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.001) + 
  ylim(0.1, 1.2) + 
  ylab("OD600")  + 
  theme(legend.title = element_blank()) +
  scale_x_discrete(breaks=c("00:00:00","24:00:00","48:00:00"), labels=c("0", "24", "48")) + ggtitle("14018 anaerobic curves: October_Nil vs. Ancestor")

print(GV14018_nil)
ExportPlot(GV14018_nil, "phenotpyePlots_04.12.24/AnAerobic_GCs/GV14018_nil", height = 5, width = 7)

ancestors <- ggplot(anc_comparison, aes(x=Time, y=Value, colour=Sample))  + 
  geom_line(aes(group=Sample)) + 
  theme_classic() + 
  scale_color_manual(name= "Population", values= Anc_colors) + 
  geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.001) + 
  ylim(0.1, 1) + 
  ylab("OD600")  + 
  theme(legend.title= element_blank()) +
  scale_x_discrete(breaks=c("00:00:00","24:00:00","48:00:00"), labels=c("0", "24", "48")) + ggtitle("Anaerobic") 

ExportPlot(ancestors, "phenotpyePlots_04.12.24/AnAerobic_GCs/ancestors", height = 5, width = 7)
```

# GV2 Anaerobic

```{r, echo = FALSE}
gc <- read_csv("2022-11-11-GV2-14018-anaerobicGrowthCurves.csv", show_col_types = FALSE)
gc <- na.omit(gc)
##format so Time is a column
rownames(gc) <- gc$Time
gc <- t(gc)
gc <- gc[-1,]
gc <- as.data.frame(gc)
gc <- tibble::rownames_to_column(gc, "Sample")
gc <- gc %>% separate(Sample, sep= "[#]", into= c("Sample", "Replicate"))

##Remove extra characters from the Replicate column
gc$Replicate <- gsub( ")", "", gc$Replicate)
gc$Sample <- sub("\\_$", "", gc$Sample)
name1 <- gc[,-c(1,2)]
list1 <- paste0(colnames(name1))
list1 <- c(list1)

##gather so that each timepoint has a separate Time value (makes it easier to plot)
lgc_onlynamed <- gather(gc, key= 'Time', value= 'Value',  colnames(name1))
lgc_onlynamed$Time <- as.factor(lgc_onlynamed$Time)
data <- lgc_onlynamed

##format data so all the columns are the correct class
data$Sample <- as.character(data$Sample)
data$Time <- as.factor(data$Time)
data$Replicate <- as.factor(data$Replicate)
data$Value <- as.numeric(data$Value)
##apply function

tgc3  <- summarySE(data, measurevar= "Value", groupvars=c("Sample", "Time")) %>%
  mutate(Type = case_when(grepl("ancestor", Sample) ~ "Ancestor", TRUE ~"Evolved"))


tgc_ordered <- tgc3 %>% 
  mutate(Time=factor(Time)) %>% 
  mutate(Time=fct_relevel(Time,c("0:00:00", "0:30:00", "1:00:00", "1:30:00", "2:00:00", "2:30:00", "3:00:00", "3:30:00", "4:00:00", "4:30:00", "5:00:00", "5:30:00", "6:00:00", "6:30:00", "7:00:00", "7:30:00", "8:00:00", "8:30:00", "9:00:00", "9:30:00"))) %>% 
  arrange(Time)

tgc_jun <- filter(tgc_ordered,   Sample %in% c("GV2_ancestor" , "GV2_June_CFS_1_P7_ ", "GV2_June_CFS_2_P7" , "GV2_June_CFS_3_P7"))

tgc_aug <- filter(tgc_ordered,   Sample %in% c("GV2_ancestor" , "GV2_Aug_CFS_1_P5_ ", "GV2_Aug_CFS_2_P5_ ", "GV2_Aug_CFS_3_P5_ " ))

tgc_nil <- filter(tgc_ordered,   Sample %in% 
                    c("GV2_ancestor" , "GV2_Aug_Nil_1_P10_ ", "GV2_Aug_Nil_2_P10_ ", "GV2_Aug_Nil_3_P10_ " ))

june_plot <- ggplot(tgc_jun, aes(x=Time, y=Value, colour=Type)) + 
  geom_line(aes(group=Sample)) + 
  theme_classic() + 
  geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.001) + ylim(0.1, 1) + 
  ylab("OD600") + 
  scale_color_manual(name= "", values= GV2_colors) +
  scale_x_discrete(breaks=c("00:00:00","24:00:00","48:00:00"), labels=c("0", "24", "48")) + ggtitle("GV2 anaerobic curves: June vs. Ancestor")

print(june_plot)
ExportPlot(june_plot, "phenotpyePlots_04.12.24/AnAerobic_GCs/GV2_june_CFS_anAerobic", height= 5, width=7)

aug_plot <- ggplot(tgc_aug, aes(x=Time, y=Value, colour=Type))  + 
  geom_line(aes(group=Sample)) + 
  theme_classic() + 
  geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.001) + 
  ylim(0.1, 1 )+ ylab("OD600") +
  scale_color_manual(name= "", values= GV2_colors) + 
  scale_x_discrete(breaks=c("00:00:00","24:00:00","48:00:00"), labels=c("0", "24", "48")) + ggtitle("GV2 anaerobic curves: August_CFS vs. Ancestor")

print(aug_plot)
ExportPlot(aug_plot, "phenotpyePlots_04.12.24/AnAerobic_GCs/GV2_aug_CFS_anAerobic", height= 5, width=7)

nil_plot <- ggplot(tgc_nil, aes(x=Time, y=Value, colour=Type)) + 
  geom_line(aes(group=Sample)) + 
  theme_classic() + 
  scale_color_manual(name= "", values= GV2_colors) +
  geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.001) + ylim(0.1, 1) + ylab("OD600")  + 
  scale_x_discrete(breaks=c("00:00:00","24:00:00","48:00:00"), labels=c("0", "24", "48")) + ggtitle("GV2 anaerobic curves: August_Nil vs. Ancestor")

print(nil_plot)
ExportPlot(nil_plot, "phenotpyePlots_04.12.24/AnAerobic_GCs/GV2_nil_anAerobic", height= 5, width=7)
```

#Graphing all trials together

```{r}
sediment_all <- read.csv("2024.03.26_GV sedimentation_all strains_manuscript edits.csv", header=T)
sediment_all$Replicate_number <- as.factor(sediment_all$Replicate_number)
sediment_all$Timepoint..min. <- as.factor(sediment_all$Timepoint..min.)
sediment_all$Type <- as.factor(sediment_all$Type)
sediment_all$Population <- as.factor(sediment_all$Population)
```

# Graphing by experiment

```{r}
##conflicting package
detach("package:Rmisc")


##filter data
June <- filter(sediment_all, Population == "06-03-GV2-SsaproCFS-1_P7" | Population == "06-03-GV2-SsaproCFS-2_P7" | Population == "06-03-GV2-SsaproCFS-3_P7" | Population == "GV2_ancestor")
Aug_CFS <- filter(sediment_all, Population == "07-30-GV2-SsaproCFS-1_P5"| Population == "07-30-GV2-SsaproCFS-2_P5" | Population == "07-30-GV2-SsaproCFS-3_P5" | Population == "GV2_ancestor")
Aug_Nil <- filter(sediment_all, Population == "07-30-GV2-Nil-1_P10"| Population == "07-30-GV2-Nil-2_P10" | Population == "07-30-GV2-Nil-3_P10" | Population == "GV2_ancestor")

Oct_CFS <- filter(sediment_all,  Population == "10-01-GV14018-CFS-1_P10" | Population == "10-01-GV14018-CFS-2_P10" | Population == "10-01-GV14018-CFS-3_P10" | Population == "GV14018_ancestor")

Oct_Nil <- filter(sediment_all, Population == "10-01-GV14018-Nil-1_P10" | Population == "10-01-GV14018-Nil-2_P10" | Population == "10-01-GV14018-Nil-3_P10" | Population == "GV14018_ancestor")

Ancestral <- filter(sediment_all, Type == "Ancestor")


##mutate data for error bars
GV2_June_CFS <- June %>% 
  group_by(Timepoint..min., Type, Population) %>% 
  mutate(sd = sd(Fract_OD), len= mean(Fract_OD))

GV2_Aug_CFS <- Aug_CFS %>% 
  group_by(Timepoint..min., Type, Population) %>% 
  mutate(sd = sd(Fract_OD), len=mean(Fract_OD))

GV2_Nil <- Aug_Nil %>% 
  group_by(Timepoint..min., Type, Population) %>% 
  mutate(sd = sd(Fract_OD), len=mean(Fract_OD))

GV14018_CFS <- Oct_CFS %>% 
  group_by(Timepoint..min., Type, Population)

GV14018_CFS <- GV14018_CFS %>%
  mutate(sd = sd(Fract_OD), len=mean(Fract_OD))

GV14018_Nil <- Oct_Nil %>% 
  group_by(Timepoint..min., Type, Population) %>% 
  mutate(sd = sd(Fract_OD), len=mean(Fract_OD))

Anc_stats <- Ancestral %>% 
  group_by(Timepoint..min.,  Population) %>% 
  mutate(sd = sd(Fract_OD), len=mean(Fract_OD))

```

Code for ancestral graph

```{r}
generate_plot <- function(df) {
  plot_title <- deparse(substitute(df))  # Get the name of the dataframe
  
  # Define color palette based on plot_title
  if (grepl("GV2", plot_title)) {
    color_palette <- GV2_colors
  } else if (grepl("GV14018", plot_title)) {
    color_palette <- GV14018_colors
  } else {
    color_palette <- NULL
  }
  
  ggplot(data = df, aes(Timepoint..min., len)) + 
    geom_line(aes(color = Type, group = Population)) +
    geom_point(aes(color = Type)) +
    geom_errorbar(aes(ymin = len - sd, ymax = len + sd, color = Type, group = Type),
                  width = 0.2) + 
    xlab("Time (min)") +
    ylab("Fraction of OD (out of 1)") +
    ggtitle("") +
    scale_y_continuous(breaks = seq(0, 1.0, 0.25)) +
    theme_minimal() +
    theme(text = element_text(size = 11), legend.title = element_blank()) +
    theme(plot.title = element_text(size = 14, hjust = 0.5, vjust = 2)) +
    scale_color_manual(values = color_palette)  # Set color palette
}


gv2_juneAgg<- generate_plot(GV2_June_CFS)
gv2_augAgg <- generate_plot(GV2_Aug_CFS)
gv2_nilAgg <- generate_plot(GV2_Nil)
gv14018_nilAgg <- generate_plot(GV14018_Nil)
gv14018_cfsAgg <- generate_plot(GV14018_CFS)

ancAgg <- ggplot(data=Anc_stats, aes(Timepoint..min., len)) + 
  geom_line(aes(color= Strain, group = Strain)) +
  geom_point(aes(color=Strain)) +
  geom_errorbar(aes(ymin= len-sd, ymax = len+sd, color=Strain, group=Strain),
                width = 0.2) + 
  xlab("Time (min)") +
  ylab("Fraction of OD (out of 1)") +  
  scale_y_continuous(breaks=seq(0,1.0, 0.25)) + 
  theme_minimal() + 
  scale_color_manual(values= Anc_colors) +
         theme(text=element_text(size= 12), legend.title = element_blank())


source("~/scripts/exportPlot.R")
ExportPlot(ancAgg, "phenotpyePlots_04.12.24//autoAggregation/ancestor_autoAggregaation", height = 5, width = 7)
ExportPlot(ancAgg, "phenotpyePlots_04.12.24/autoAggregation/ancestor_autoAggregaation", height = 5, width = 7)
ExportPlot(gv2_juneAgg, "phenotpyePlots_04.12.24/autoAggregation/gv2_june_cfs_autoAggregaation", height = 5, width = 7)
ExportPlot(gv2_augAgg, "phenotpyePlots_04.12.24/autoAggregation/gv2_aug_cfs_autoAggregaation", height = 5, width = 7)
ExportPlot(gv2_nilAgg, "phenotpyePlots_04.12.24/autoAggregation/gv2_nil_autoAggregation", height = 5, width = 7)
ExportPlot(gv14018_nilAgg, "phenotpyePlots_04.12.24//autoAggregation/gv14018_nil_autoAggregaation", height = 5, width = 7)
ExportPlot(gv14018_cfsAgg, "phenotpyePlots_04.12.24/autoAggregation/gv14018_cfs_autoAggregaation", height = 5, width = 7)

ggarrange(gv2_nilAgg, gv14018_nilAgg, ncol = 1, nrow =2)
```


```{r}
##Read in CSV file with growth curve data
gc <- read_csv("2023.01.26_anaerobic GCs_14018 evolved + CFS_Gardnerella species.csv", col_names = TRUE, cols(Time = col_factor()))
gc <- na.omit(gc)

##format so Time is a column
rownames(gc) <- gc$Time
gc <- t(gc)
gc <- gc[-1,]
gc <- as.data.frame(gc)
gc <- tibble::rownames_to_column(gc, "Sample")
gc <- separate(gc, Sample, sep= "[#]", into= c("Sample", "Replicate"))

##Remove extra characters from the Sample column
gc$Sample <- sub("\\_$", "", gc$Sample)
name1 <- gc[,-c(1,2)]
list1 <- paste0(colnames(name1))
list1 <- c(list1)

##gather so that each timepoint has a separate Time value (makes it easier to plot)
lgc_onlynamed <- gather(gc, key= 'Time', value= 'Value',  colnames(name1))
lgc_onlynamed$Time <- as.factor(lgc_onlynamed$Time)
data <- lgc_onlynamed

##format data so all the columns are the correct class
data$Sample <- as.character(data$Sample)
data$Time <- as.factor(data$Time)
data$Replicate <- as.factor(data$Replicate)
data$Value <- as.numeric(data$Value)
##apply function

tgc <- summarySE(data, measurevar= "Value", groupvars=c("Sample", "Time"))
tgc_ordered <- tgc %>% 
  mutate(Time=factor(Time)) %>% 
  mutate(Time=fct_relevel(Time,c("0:00:00", "0:30:00", "1:00:00", "1:30:00", "2:00:00", "2:30:00", "3:00:00", "3:30:00", "4:00:00", "4:30:00", "5:00:00", "5:30:00", "6:00:00", "6:30:00", "7:00:00", "7:30:00", "8:00:00", "8:30:00", "9:00:00", "9:30:00"))) %>% arrange(Time)

GV14018_CFS <- filter(tgc_ordered, Sample %in% c("GV14018_ancestor_CFS", "GV14018_ancestor_no CFS", "GV14018_evolved_CFS_S1", "GV14018_evolved_CFS_S2", "GV14018_evolved_CFS_S3")) 
Ancestors <- filter(tgc_ordered, Sample %in% c("GV14018_ancestor_CFS", "GV14018_ancestor_no CFS"))

graph_CFS <- ggplot(GV14018_CFS, aes(x=Time, y=Value, colour=Sample))  + geom_line(aes(group=Sample)) + theme_classic() + 
  scale_color_manual(name= "Population", values= c("red", GV14018_colors, "#F4BEBA", "#F4BEBA"), labels = c("Ancestor - with CFS", "Ancestor - no CFS", "Evolved", "", "")) + geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.001) + ylim(0.1, 1.2) + ylab("OD600")  + 
  scale_x_discrete(breaks=c("00:00:00","24:00:00","48:00:00"), labels=c("0", "24", "48")) + ggtitle("")

print(graph_CFS)

graph_anc <- ggplot(Ancestors, aes(x=Time, y=Value, colour=Sample))  + geom_line(aes(group=Sample)) + theme_classic() + 
  scale_color_manual(name= "Population", values= c("red", GV14018_colors, "#F4BEBA", "#F4BEBA"), labels = c("Ancestor - with CFS", "Ancestor - no CFS", "Evolved", "Evolved", "Evolved")) + geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.001) + ylim(0.1, 1.2) + ylab("OD600")  + 
  scale_x_discrete(breaks=c("00:00:00","24:00:00","48:00:00"), labels=c("0", "24", "48")) + ggtitle("")
print(graph_anc)
```